---
title: "MuM_PreDiCT: Connect to SQL Databases & Extract Data"
output: html_notebook
---


last updated: 2023.02.24 (Lisa Kent)

This notebook:
* reads in raw NIMATS tables from database
* explores meta data
* NOTHING written to csv

```{r}
if(!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
```

__Lode odbc (Open Database Connectivity) package__
```{r}
if(!require(odbc)) install.packages("odbc")
library(odbc)
```

__Declare connection to database__
```{r}
con <- dbConnect(odbc::odbc(), "hirusql3", timeout = 10)
```


__Read in data from database and check__
__Deaths Data__
```{r}
deaths <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[Death_Data_SeRP_1]")
```

```{r}
head(deaths)
```

```{r}
glimpse(deaths)
```


```{r}
n_distinct(deaths$study_id)
```

```{r}
deaths %>% group_by(death_year, death_month) %>% count()
```



__Read in data from database and check__
__Prescribing Data__
```{r}
#likely to be very large!!!
epd <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[EPD_1]")
```


```{r}
head(epd)
```

```{r}
glimpse(epd)
#there does not appear to be any issue with trailing white space for any variable
```



```{r}
n_distinct(epd$study_id)
```

```{r}
epd %>% group_by(ISSUE_DATE) %>% count()

#numbers seem reasonable from Jan2010 onwards
#pattern observed where weekends and holidays have reduced number of prescriptions issued
```

```{r}
epd %>% 
  filter(ISSUE_DATE > "2020-02-04 00:00:00") %>% group_by(ISSUE_DATE) %>% count()

#invalid dates: [EXAMPLES REMOVED FOR SDC]
```


```{r}
sum(is.na(epd$ISSUE_DATE))
#missing values present for issue date
```

```{r}
epd %>% group_by(SCAN_YEAR, SCAN_MONTH) %>% count()

# Earliest SCAN_YEAR & SCAN_MONTH: [EXAMPLES REMOVED FOR SDC], latest: [EXAMPLES REMOVED FOR SDC]
# Check if numbers are as expected for earliest and latest months
# [EXAMPLES REMOVED FOR SDC]

```


```{r}
sum(is.na(epd$SCAN_YEAR))
sum(is.na(epd$SCAN_MONTH))

```


```{r}
# check for missing values
sum(is.na(epd$vmp_nm))
sum(is.na(epd$vtm_nm))
sum(is.na(epd$quantity))
sum(is.na(epd$presentation))
sum(is.na(epd$strength))


sum(is.na(epd$bnf_chapter))
sum(is.na(epd$bnf_section))
sum(is.na(epd$bnf_paragraph))
sum(is.na(epd$bnf_sub_paragraph))

```

```{r}
epd %>% group_by(bnf_chapter, bnf_section, bnf_paragraph, bnf_sub_paragraph) %>% count()

```

__Admissions Data__
```{r}
pas <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[PAS_1]")
```

```{r}
colnames(pas)
```

```{r}
glimpse(pas)
#there is trailing white space for some variables
```

```{r}
pas <- pas %>% 
  mutate(across(where(is.character), str_trim))

glimpse(pas)
```

```{r}
head(pas, 20)
```

```{r}
n_distinct(pas$study_id)
```





__NIMATS General Data__
```{r}
nimats_gen <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[General_Data_v2_1]")
```

```{r}
colnames(nimats_gen)
```



```{r}
n_distinct(nimats_gen$STUDY_ID)
n_distinct(nimats_gen$CHILD_ID)
```

```{r}
n_distinct(nimats_gen$STUDY_ID, nimats_gen$CHILD_ID)
sum(is.na(nimats_gen$STUDY_ID))
sum(is.na(nimats_gen$CHILD_ID))
#there are more unique mum-child combinations that unique children - due to NA's in CHILD_ID
```


```{r}
n_distinct(nimats_gen$STUDY_ID, nimats_gen$DOB_YEAR, nimats_gen$DOB_MONTH)
```


```{r}
n_distinct(nimats_gen$STUDY_ID, nimats_gen$CHILD_ID) - n_distinct(nimats_gen$CHILD_ID)
```

```{r}
nimats_gen %>% group_by(DOB_YEAR, DOB_MONTH) %>% count()
```

```{r}
head(nimats_gen, 50)
```

```{r}
sum(is.na(nimats_gen$LMP))
sum(is.na(nimats_gen$ACTUAL_EDC))
```


```{r}
nimats_gen %>% group_by(AGE_AT_BIRTH) %>% count()
```



```{r}
nimats_gen %>% group_by(DOB_YEAR, BIRTH_STATUS) %>% count()
```


__NIMATS Delivery Details Data__
```{r}
nimats_dd <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[Delivery_Details_v2_1]")
```

```{r}
colnames(nimats_dd)
```

```{r}
glimpse(nimats_dd)
```



```{r}
n_distinct(nimats_dd$STUDY_ID)
n_distinct(nimats_dd$CHILD_ID)
```


```{r}
n_distinct(nimats_dd$STUDY_ID, nimats_dd$CHILD_ID)
#there are more unique mum-child combinations that unique children 
```

```{r}
sum(is.na(nimats_dd$STUDY_ID))
sum(is.na(nimats_dd$CHILD_ID))
#as with gen data, missing child_id's likely causing this issue
```

```{r}
nimats_dd %>% group_by(DOB_YEAR, DOB_MONTH) %>% count()
```


```{r}
head(nimats_dd, 50)
```


__check if descriptions and codes match as per NIMATS metadata__
```{r}
nimats_dd %>% group_by(DOB_YEAR, SURGERY_CODE, SURGERY_DESC) %>% count()
```

```{r}
nimats_dd %>% group_by(PROBLEMS_AFFECTING_LABOUR_CODE, problems_affecting_labour_desc) %>% count()
```


```{r}
nimats_dd %>% group_by(PRESENT_PREG_PROBLEMS_CODE, present_preg_problems_desc) %>% count()
```

```{r}
nimats_dd %>% group_by(ANTENATAL_STEROIDS, antenatal_steroids_desc) %>% count()
```

```{r}
nimats_dd %>% group_by(DOB_YEAR, DOB_MONTH, antenatal_steroids_desc) %>% count()
```


__NIMATS Previous Medical History Data__
```{r}
nimats_pmh <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[Prev_Medical_History_1]")
```

```{r}
glimpse(nimats_pmh)
#trailing white space in character variables
```

```{r}
colnames(nimats_pmh)
```

```{r}
n_distinct(nimats_pmh$STUDY_ID)
n_distinct(nimats_pmh$CHILD_ID)
```


```{r}
n_distinct(nimats_pmh$STUDY_ID, nimats_pmh$CHILD_ID)
#there are more unique mum-child combinations that unique children 
```

```{r}
sum(is.na(nimats_pmh$STUDY_ID))
sum(is.na(nimats_pmh$CHILD_ID))
#as with gen data, missing child_id's likely causing this issue
```

```{r}
nimats_pmh %>% group_by(DOB_YEAR, DOB_MONTH) %>% count()
```


```{r}
nimats_pmh %>% group_by(BLOOD_DISORDERS_CODE, BLOOD_DISORDERS_DESC) %>% count()

nimats_pmh %>% group_by(CARDIOVASCULAR_CODE, CARDIOVASCULAR_DESC) %>% count()

nimats_pmh %>% group_by(CNS_CODE, CNS_DESC) %>% count()

nimats_pmh %>% group_by(DIABETES_CODE, DIABETES_DESC) %>% count()

nimats_pmh %>% group_by(GENERAL_SURGERY_CODE, GENERAL_SURGERY_DESC) %>% count()

nimats_pmh %>% group_by(GIT_CODE, GIT_DESC) %>% count()

nimats_pmh %>% group_by(GU_INFECTION_CODE, GU_INFECTION_DESC) %>% count()

nimats_pmh %>% group_by(GYNAE_CODE, GYNAE_DESC) %>% count()

nimats_pmh %>% group_by(INFECTION_CODE, INFECTION_DESC) %>% count()

nimats_pmh %>% group_by(INFERTILITY_CODE, INFERTILITY_DESC) %>% count()

nimats_pmh %>% group_by(MALIGNANCY_CODE, MALIGNANCY_DESC) %>% count()

nimats_pmh %>% group_by(MENTAL_HEALTH_CODE, MENTAL_HEALTH_DESC) %>% count()

nimats_pmh %>% group_by(MENTAL_HEALTH_MINOR_CODE, MENTAL_HEALTH_MINOR_DESC) %>% count()

nimats_pmh %>% group_by(METABOLIC_CODE, METABOLIC_DESC) %>% count()

nimats_pmh %>% group_by(MUSCULO_SKELETAL_CODE, MUSCULO_SKELETAL_DESC) %>% count()

nimats_pmh %>% group_by(RENAL_CODE, RENAL_DESC) %>% count()

nimats_pmh %>% group_by(RESPIRATORY_CODE, RESPIRATORY_DESC) %>% count()

nimats_pmh %>% group_by(SKIN_CODE, SKIN_DESC) %>% count()


```



__NIMATS Specified Condition (PMHx Free Text) Data__
```{r}
nimats_cond <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[Specified_Condition_1]")
```

```{r}
glimpse(nimats_cond)
#trailing white space in character variables
```

```{r}
n_distinct(nimats_cond$study_id)
n_distinct(nimats_cond$child_id)
sum(is.na(nimats_cond$child_id))
sum(is.na(nimats_cond$AUTOIMMUNE_SKIN))
```

```{r}
summary(nimats_cond)
```


```{r}
colnames(nimats_cond)
```




__NIMATS Previous Obstetric History Data__
```{r}
nimats_obshx <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[Prev_Obstetric_History_1]")
```

```{r}
glimpse(nimats_obshx)
#trailing white space in character variables
```

```{r}
colnames(nimats_obshx)
```

```{r}
n_distinct(nimats_obshx$STUDY_ID)
n_distinct(nimats_obshx$STUDY_ID, nimats_obshx$DOB_MONTH, nimats_obshx$DOB_YEAR)
n_distinct(nimats_obshx$CHILD_ID)
sum(is.na(nimats_obshx$CHILD_ID))
```


```{r}
nimats_obshx %>% group_by(OHR_CONDITION, OHR_CODE, OHR_DESCRIPTION) %>% count()
```

