---
title: "MuM_PreDiCT: Deduplicate NIMATS delivery details data at pregnancy level"
output: html_notebook
---

last updated: 2023.02.08 (Lisa Kent)

This notebook:
* reads in raw NIMATS delivery details table from database
* explores and de-duplicates each pregnancy level variable (dropping infant level variables)
* builds a pregnancy level dataset
* final dataset written to csv

```{r}
if(!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
```

__Lode odbc (Open Database Connectivity) package__
```{r}
library(odbc)
```

__Declare connection to database__
```{r}
con <- dbConnect(odbc::odbc(), "hirusql3", timeout = 10)
```



__NIMATS Delivery Details__
```{r}
nimats_dd <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[Delivery_Details_v2_1]")
```


```{r}
head(nimats_dd, 50)
```


__TRIM WHITE SPACE__
```{r}
nimats_dd <- nimats_dd %>%
  mutate(across(everything(), str_trim))
```

__FIX NA VALUES__
```{r}
nimats_dd[nimats_dd =="<NA>"] <- NA
```

__COERCE DATES TO CORRECT FORMAT__
```{r}
nimats_dd <- nimats_dd %>%
  mutate(DOB_MONTH = case_when(
    DOB_MONTH == "JANUARY" ~ "01",
    DOB_MONTH == "FEBRUARY" ~ "02",
    DOB_MONTH == "MARCH" ~ "03",
    DOB_MONTH == "APRIL" ~ "04",
    DOB_MONTH == "MAY" ~ "05",
    DOB_MONTH == "JUNE" ~ "06",
    DOB_MONTH == "JULY" ~ "07",
    DOB_MONTH == "AUGUST" ~ "08",
    DOB_MONTH == "SEPTEMBER" ~ "09",
    DOB_MONTH == "OCTOBER" ~ "10",
    DOB_MONTH == "NOVEMBER" ~ "11",
    DOB_MONTH == "DECEMBER" ~ "12",
    
  ))
```


__CREATE PREGNANCY LEVEL IDENTIFIER__
```{r}
nimats_dd <- nimats_dd %>%
  unite(PREG_ID,
        c("STUDY_ID", "DOB_YEAR", "DOB_MONTH"),
        remove = F,
        sep="_")
  
```


__CREATE PREGNANCY LEVEL COHORT SPINE__

```{r}
# variables selected one by one to check if any causing duplicates
nimats_dd_preg_level <- nimats_dd %>%
  select(PREG_ID, DOB_MONTH, DOB_YEAR) %>%
  distinct()
```


```{r}
#sanity check (should all be same number)
n_distinct(nimats_dd$STUDY_ID, nimats_dd$DOB_MONTH, nimats_dd$DOB_YEAR)
n_distinct(nimats_dd$PREG_ID)
nrow(nimats_dd_preg_level)
```
```{r}
colnames(nimats_dd)
```



__DEDUPLICATION OF PREG LEVEL DATA__

__SURGERY__
```{r}
nimats_dd_preg_level_surgery <- nimats_dd %>%
  select(PREG_ID, SURGERY_CODE, SURGERY_DESC) %>%
  distinct()
# no additional rows 
```


```{r}
nimats_dd_preg_level <- left_join(nimats_dd_preg_level, nimats_dd_preg_level_surgery, by="PREG_ID")
```


__PROBLEMS AFFECTING LABOUR__
```{r}
nimats_dd_preg_level_pal <- nimats_dd %>%
  select(PREG_ID, PROBLEMS_AFFECTING_LABOUR_CODE, problems_affecting_labour_desc) %>%
  distinct()
# 349 additional rows added
```

```{r}
n_occur <- data.frame(table(nimats_dd_preg_level_pal$PREG_ID))
duplicates <- nimats_dd_preg_level_pal[nimats_dd_preg_level_pal$PREG_ID %in% n_occur$Var1[n_occur$Freq>1],]
duplicates[order(duplicates$PREG_ID),]
# visual inspection - several values possible in same pregnancy
# 
# solution = dummy variables
```

```{r}
nimats_dd_preg_level_pal %>% group_by(PROBLEMS_AFFECTING_LABOUR_CODE, problems_affecting_labour_desc) %>% count()
# code matches desc therefore just use desc
```

```{r}
nimats_dd_preg_level_pal$problems_affecting_labour_desc <- gsub(" ", "_", nimats_dd_preg_level_pal$problems_affecting_labour_desc)
 
nimats_dd_preg_level_pal %>% group_by(problems_affecting_labour_desc) %>% count()
```


```{r}
nimats_dd_preg_level_pal_summary <- nimats_dd_preg_level_pal %>%
  select(PREG_ID, problems_affecting_labour_desc) %>%
  group_by(PREG_ID, problems_affecting_labour_desc) %>%
  summarise(COUNT = n())
```


```{r}
nimats_dd_preg_level_pal_summary$COUNT[nimats_dd_preg_level_pal_summary$COUNT>0] <- 1
```

```{r}
nimats_dd_preg_level_pal_wide <- nimats_dd_preg_level_pal_summary %>%
  pivot_wider(names_from = problems_affecting_labour_desc,
              values_from = COUNT,
              values_fill = list(COUNT = 0))
```

```{r}
colnames(nimats_dd_preg_level_pal_wide)[2:ncol(nimats_dd_preg_level_pal_wide)] <- paste0("PROB_AFF_LABOUR_", colnames(nimats_dd_preg_level_pal_wide)[2:ncol(nimats_dd_preg_level_pal_wide)])
head(nimats_dd_preg_level_pal_wide)
```


```{r}
nimats_dd_preg_level <- left_join(nimats_dd_preg_level, nimats_dd_preg_level_pal_wide, by="PREG_ID")
```



__PRESENT PREGNANCY PROBLEMS__
```{r}
nimats_dd_preg_level_ppp <- nimats_dd %>%
  select(PREG_ID, PRESENT_PREG_PROBLEMS_CODE, present_preg_problems_desc) %>%
  distinct()
# 46,558 additional rows added
```

```{r}
n_occur <- data.frame(table(nimats_dd_preg_level_ppp$PREG_ID))
duplicates <- nimats_dd_preg_level_ppp[nimats_dd_preg_level_ppp$PREG_ID %in% n_occur$Var1[n_occur$Freq>1],]
duplicates[order(duplicates$PREG_ID),]
# visual inspection - several values possible in same pregnancy
# 
# solution = dummy variables
```

```{r}
nimats_dd_preg_level_ppp %>% group_by(PRESENT_PREG_PROBLEMS_CODE, present_preg_problems_desc) %>% count()
# code matches desc therefore just use desc
```


```{r}
nimats_dd_preg_level_ppp$present_preg_problems_desc <- gsub(" ", "_", nimats_dd_preg_level_ppp$present_preg_problems_desc)
 
nimats_dd_preg_level_ppp %>% group_by(present_preg_problems_desc) %>% count()
```


```{r}
nimats_dd_preg_level_ppp_summary <- nimats_dd_preg_level_ppp %>%
  select(PREG_ID, present_preg_problems_desc) %>%
  group_by(PREG_ID, present_preg_problems_desc) %>%
  summarise(COUNT = n())
```


```{r}
nimats_dd_preg_level_ppp_summary$COUNT[nimats_dd_preg_level_ppp_summary$COUNT>0] <- 1
```

```{r}
nimats_dd_preg_level_ppp_wide <- nimats_dd_preg_level_ppp_summary %>%
  pivot_wider(names_from = present_preg_problems_desc,
              values_from = COUNT,
              values_fill = list(COUNT = 0))
```

```{r}
colnames(nimats_dd_preg_level_ppp_wide)[2:ncol(nimats_dd_preg_level_ppp_wide)] <- paste0("PRES_PREG_PROB_", colnames(nimats_dd_preg_level_ppp_wide)[2:ncol(nimats_dd_preg_level_ppp_wide)])
head(nimats_dd_preg_level_ppp_wide)
```


```{r}
nimats_dd_preg_level <- left_join(nimats_dd_preg_level, nimats_dd_preg_level_ppp_wide, by="PREG_ID")
```


__ANTENATAL STEROIDS__
```{r}
nimats_dd_preg_level_steroids <- nimats_dd %>%
  select(PREG_ID, ANTENATAL_STEROIDS, antenatal_steroids_desc) %>%
  distinct()

```


```{r}
n_occur <- data.frame(table(nimats_dd_preg_level_steroids$PREG_ID))
duplicates <- nimats_dd_preg_level_steroids[nimats_dd_preg_level_steroids$PREG_ID %in% n_occur$Var1[n_occur$Freq>1],]
duplicates[order(duplicates$PREG_ID),]
# visual inspection - duplicates mainly have differing values values
# is it possible that steroids could be given both <48h and >48h from birth
# could a second dose be given?
# solution = dummy variables (with option to contract again at later stage)
```

```{r}
nimats_dd_preg_level_steroids %>% group_by(ANTENATAL_STEROIDS, antenatal_steroids_desc) %>% count()
# code matches desc therefore just use desc
```


```{r}
nimats_dd_preg_level_steroids$antenatal_steroids_desc <- gsub(" ", "", nimats_dd_preg_level_steroids$antenatal_steroids_desc)
 
nimats_dd_preg_level_steroids %>% group_by(antenatal_steroids_desc) %>% count()
```


```{r}
nimats_dd_preg_level_steroids_summary <- nimats_dd_preg_level_steroids %>%
  select(PREG_ID, antenatal_steroids_desc) %>%
  group_by(PREG_ID, antenatal_steroids_desc) %>%
  summarise(COUNT = n())
```


```{r}
nimats_dd_preg_level_steroids_summary$COUNT[nimats_dd_preg_level_steroids_summary$COUNT>0] <- 1
```

```{r}
nimats_dd_preg_level_steroids_wide <- nimats_dd_preg_level_steroids_summary %>%
  pivot_wider(names_from = antenatal_steroids_desc,
              values_from = COUNT,
              values_fill = list(COUNT = 0))
```

```{r}
colnames(nimats_dd_preg_level_steroids_wide)[2:ncol(nimats_dd_preg_level_steroids_wide)] <- paste0("ANTENATAL_STEROID_", colnames(nimats_dd_preg_level_steroids_wide)[2:ncol(nimats_dd_preg_level_steroids_wide)])
head(nimats_dd_preg_level_steroids_wide)
```


```{r}
nimats_dd_preg_level <- left_join(nimats_dd_preg_level, nimats_dd_preg_level_steroids_wide, by="PREG_ID")
```





__WRITE TO CSV__
```{r}
write_csv(nimats_dd_preg_level, file=paste0("S:/HBS_E062/3. DATA/nimats_dd_preg_level_", format(lubridate::now(), "%Y.%m.%d_%H.%M"), ".csv"))
```




######################################
#######           END          #######
######################################


