---
title: "MuM_PreDiCT: Deduplicate NIMATS obstetric history"
output: html_notebook
---

last updated: 2023.02.08 (Lisa Kent)

This notebook:
* reads in raw NIMATS previous obstetric history table from database
* explores and de-duplicates each variable (all pregnancy level variables)
* builds a pregnancy level dataset
* final dataset written to csv

```{r}
if(!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
```

__Lode odbc (Open Database Connectivity) package__
```{r}
library(odbc)
```

__Declare connection to database__
```{r}
con <- dbConnect(odbc::odbc(), "hirusql3", timeout = 10)
```



__NIMATS OBSTETRIC HISTORY__
```{r}
nimats_obshx <- dbGetQuery(con,"Select * from [HDRNIHBS_E062].[Prev_Obstetric_History_1]")
```


```{r}
head(nimats_obshx, 50)
```


__TRIM WHITE SPACE__
```{r}
nimats_obshx <- nimats_obshx %>%
  mutate(across(everything(), str_trim))
```

__FIX NA VALUES__
```{r}
nimats_obshx[nimats_obshx =="<NA>"] <- NA
```

__COERCE DATES TO CORRECT FORMAT__
```{r}
nimats_obshx <- nimats_obshx %>%
  mutate(DOB_MONTH = case_when(
    DOB_MONTH == "JANUARY" ~ "01",
    DOB_MONTH == "FEBRUARY" ~ "02",
    DOB_MONTH == "MARCH" ~ "03",
    DOB_MONTH == "APRIL" ~ "04",
    DOB_MONTH == "MAY" ~ "05",
    DOB_MONTH == "JUNE" ~ "06",
    DOB_MONTH == "JULY" ~ "07",
    DOB_MONTH == "AUGUST" ~ "08",
    DOB_MONTH == "SEPTEMBER" ~ "09",
    DOB_MONTH == "OCTOBER" ~ "10",
    DOB_MONTH == "NOVEMBER" ~ "11",
    DOB_MONTH == "DECEMBER" ~ "12",
    
  ))
```


__CREATE PREGNANCY LEVEL IDENTIFIER__
```{r}
nimats_obshx <- nimats_obshx %>%
  unite(PREG_ID,
        c("STUDY_ID", "DOB_YEAR", "DOB_MONTH"),
        remove = F,
        sep="_")

head(nimats_obshx)
  
```



__DEDUPLICATION OF PREG LEVEL DATA__
```{r}
# variables selected one by one to check if any causing duplicates
nimats_obshx_preg_level <- nimats_obshx %>%
  select(PREG_ID, DOB_MONTH, DOB_YEAR) %>%
  distinct()

```

```{r}
# sanity check (should all be same number)
n_distinct(nimats_obshx$STUDY_ID, nimats_obshx$DOB_MONTH, nimats_obshx$DOB_YEAR)
n_distinct(nimats_obshx$PREG_ID) 
nrow(nimats_obshx_preg_level)
```

```{r}
colnames(nimats_obshx)
```

```{r}
nimats_obshx <- nimats_obshx %>%
  mutate(#OHR_CONDITION = paste0("OHR_", OHR_CONDITION),
         OHR_COND_DESC = paste0("OHR_", OHR_CONDITION, "_", OHR_DESCRIPTION))

head(nimats_obshx, 50)
```

```{r}
nimats_obshx$OHR_COND_DESC[nimats_obshx$OHR_COND_DESC == "OHR_NA_NA"] <- NA


nimats_obshx$OHR_COND_DESC <- gsub(" ", "_", nimats_obshx$OHR_COND_DESC)
 
nimats_obshx %>% group_by(OHR_COND_DESC) %>% count()

head(nimats_obshx, 50)
```



__OHR CONDITION AND DESCRIPTION__
```{r}
nimats_obshx_preg_level_ohr <- nimats_obshx %>%
  select(PREG_ID, OHR_COND_DESC) %>%
  distinct()
```


```{r}
nimats_obshx_preg_level_ohr_summary <- nimats_obshx_preg_level_ohr %>%
  select(PREG_ID, OHR_COND_DESC) %>%
  group_by(PREG_ID, OHR_COND_DESC) %>%
  summarise(COUNT = n())
```


```{r}
nimats_obshx_preg_level_ohr_summary$COUNT[nimats_obshx_preg_level_ohr_summary$COUNT>0] <- 1
```

```{r}
nimats_obshx_preg_level_ohr_summary_wide <- nimats_obshx_preg_level_ohr_summary %>%
  pivot_wider(names_from = OHR_COND_DESC,
              values_from = COUNT,
              values_fill = list(COUNT = 0))
```


```{r}
head(nimats_obshx_preg_level_ohr_summary_wide, 50)
```


__WRITE TO CSV__
```{r}
write_csv(nimats_obshx_preg_level_ohr_summary_wide, file=paste0("S:/HBS_E062/3. DATA/nimats_obshx_preg_level_", format(lubridate::now(), "%Y.%m.%d_%H.%M"), ".csv"))
```




######################################
#######           END          #######
######################################
